Evaluation of classical correlation functions from 2/3D images on CPU and GPU architectures: introducing CorrelationFunctions.jl

### 2022-05-31 paper-correlations.tex
fill map section

\documentclass[reprint,amsmath,amssymb,aps,pre,nofootinbib]{revtex4-1}
% Latex
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
% Xelatex
%\usepackage{polyglossia}
%\usepackage{fontspec}
%\setdefaultlanguage{english}
%\setmainfont{DejaVu Sans}
%\setsansfont{DejaVu Sans}
%\setmonofont{DejaVu Sans Mono}
\usepackage{bm}
\usepackage{cleveref}
\usepackage{xcolor}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{subfigure}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.9\linewidth]{images/corr.png}
  \caption[]{WTF?}
  \label{fig:functions}
\end{figure}

@@

Correlation functions can be computed using periodic and zero-padding boundary
conditions. With periodic boundary conditions a correlation function is
calculated as if the input array were periodically extended to infinity,
i.e. $A_k = A_{k+N_k}$ where $N_k$ is the k-th array dimension. With
zero-padding the input is extended with zeros, i.e. $A_k = 0$ when
$0 < k < N_k$.

@@

The probability $p(\bm{r})$ can be expressed as
\begin{equation*}
  p(\bm{r}) = \frac{\text{Number of successful trials $S(\bm{r})$}}
  {\text{Total number of trials $N(\bm{r})$}}
\end{equation*}

@@

If $\chi(\bm{x_1}, \bm{x_2})$ can be expressed as
$\chi(\bm{x_1}, \bm{x_2}) = \chi_1(\bm{x_1})\chi_2(\bm{x_2})$ (what we call
\code{SeparableIndicator} in our package) we use the following algorithm to
calculate the array \code{success} for one-dimensional slice $a$: \footnote{From
here on $\odot$ denotes concatenation and $:$ denotes consing (joining an
element in front of a list).}
\begin{algorithmic}[1]
  \Procedure{s2sep}{$a, \chi_1, \chi_2, periodic$}
  \If{$!periodic$}
    \State $a \gets a \odot [0 \dots 0]$ \Comment{Pad with zeros to twice the
      length of $a$.}
  \EndIf
  \State $a_1 \gets \chi_1(a)$ \Comment{apply $\chi_1$ element-wise to $a$}
  \State $a_2 \gets \chi_2(a)$ \Comment{apply $\chi_2$ element-wise to $a$}
  \State $f_1 \gets rfft(a_1)$ \Comment{$rfft$ is fast Fourier transform for
    real-valued data}
  \State $f_2 \gets rfft(a_2)$
  \State $f_{S_2} \gets f_1 \cdot \overline{f_2}$ \Comment{multiply element-wise}
  \State $success \gets irfft(f_{S_2})$ \Comment{$irfft$ is inverse of $rfft$}
  \State \textbf{return} $success$
  \EndProcedure
\end{algorithmic}

If $\chi(\bm{x})$ cannot be expressed in this way (what we call
\code{InseparableIndicator}) the algorithm is slightly different:
\begin{algorithmic}[1]
  \Procedure{s2insep}{$a, \chi, periodic$}
    \State $len \gets length(a)$
    \If{$periodic$}
      \State $a \gets a \odot a$
    \EndIf
    \State $success \gets zeros(len)$
    \For{$shift \gets 0,len-1$}
      \State $a_2 \gets a[shift \dots]$
      \Comment{Take elements with index bigger than $shift$.}
      \State $success[shift] \gets sum(\chi(a, a_2))$
      \Comment{Apply $\chi$ element-wise and take a sum.}
    \EndFor
    \State \textbf{return} $success$
  \EndProcedure
\end{algorithmic}

@@

For computation of cluster function we use \code{label\_components} function
from \code{Image.jl} package which extracts connected comonents (clusters) from the
binary array. Each cluster gets its own unique integer identificator. The
identificator $0$ is reserved for the void phase.

The algorithm for calculation of the array \code{success} for
$C_2^{(i)}(\bm{r})$ for one-dimensional slice $a$ is straightforward:
\begin{algorithmic}[1]
  \Procedure{c2}{$a, phase, periodic$}
  \State $a_1 \gets x = phase$ \Comment{$a$ is compared with $phase$
    element-wise}
  \State $c \gets label\_components(a_1)$
  \State $\chi(x,y) \gets (x = y \ne 0)$
  \State \textbf{return} $s2insep(c, \chi, periodic)$
  \EndProcedure
\end{algorithmic}

@@

\section{The map approach.}
\label{mapsec}
Text

@@

\begin{figure*}[t]
  \centering
  \subfigure[Lineal-path correlation function]{
    \includegraphics[width=0.475\linewidth]{images/l2-2d.png}
    \label{fig:l2-2d}}
  \hfill
  \subfigure[Two-point correlation function]{
    \includegraphics[width=0.475\linewidth]{images/s2-2d.png}
    \label{fig:s2-2d}}
  \vskip\baselineskip
  \subfigure[Surface-surface correlation function]{
    \includegraphics[width=0.475\linewidth]{images/ss-2d.png}
    \label{fig:ss-2d}}
  \hfill
  \subfigure[Surface-void correlation function]{
    \includegraphics[width=0.475\linewidth]{images/sv-2d.png}
    \label{fig:sv-2d}}
  \caption[]{A comparison of calculated values of $L_2(r)$, $S_2(r)$,
    $F_{ss}(r)$ and $F_{sv}(r)$ with theoretical values for overlapping disks.}
  \label{fig:2d}
\end{figure*}

\begin{figure*}[t]
  \centering
  \subfigure[Lineal-path correlation function]{
    \includegraphics[width=0.475\linewidth]{images/l2-3d.png}
    \label{fig:l2-3d}}
  \hfill
  \subfigure[Two-point correlation function]{
    \includegraphics[width=0.475\linewidth]{images/s2-3d.png}
    \label{fig:s3-3d}}
  \vskip\baselineskip
  \subfigure[Surface-surface correlation function]{
    \includegraphics[width=0.475\linewidth]{images/ss-3d.png}
    \label{fig:ss-3d}}
  \hfill
  \subfigure[Surface-void correlation function]{
    \includegraphics[width=0.475\linewidth]{images/sv-3d.png}
    \label{fig:sv-3d}}
  \caption[]{A comparison of calculated values of $L_2(r)$, $S_2(r)$,
    $F_{ss}(r)$ and $F_{sv}(r)$ with theoretical values for overlapping balls.}
  \label{fig:3d}
\end{figure*}

\begin{figure*}[t]
  \centering
  \subfigure[Pore size correlation function (2D)]{
    \includegraphics[width=0.475\linewidth]{images/ps-2d.png}
    \label{fig:ps-2d}}
  \hfill
  \subfigure[Chord length correlation function (2D)]{
    \includegraphics[width=0.475\linewidth]{images/cl-2d.png}
    \label{fig:cl-2d}}
  \vskip\baselineskip
  \subfigure[Pore size correlation function (3D)]{
    \includegraphics[width=0.475\linewidth]{images/ps-3d.png}
    \label{fig:ps-3d}}
  \hfill
  \subfigure[Chord length correlation function (3D)]{
    \includegraphics[width=0.475\linewidth]{images/cl-3d.png}
    \label{fig:cl-3d}}
  \caption[]{A comparison of calculated values of $P(r)$ and $p(r)$
    with theoretical values for overlapping disks (\cref{fig:ps-2d} and
    \cref{fig:cl-2d}) and balls(\cref{fig:ps-3d} and \cref{fig:cl-3d}).}
  \label{fig:pscl}
\end{figure*}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.9\linewidth]{images/timing-image.png}
  \caption[]{An example of 2D image used in timing measurements.}
  \label{fig:valuenoise}
\end{figure}

\begin{figure*}[t]
  \centering
  \subfigure[Two-dimensional case, module \code{Directional}]{
    \includegraphics[width=0.475\linewidth]{images/time-2d.png}
    \label{fig:timings-2d}}
  \hfill
  \subfigure[Three-dimensional case, module \code{Directional}]{
    \includegraphics[width=0.475\linewidth]{images/time-3d.png}
    \label{fig:timings-3d}}
  \vskip\baselineskip
  \subfigure[Two-dimensional case, module \code{Map}, CPU]{
    \includegraphics[width=0.475\linewidth]{images/time-2d-map.png}
    \label{fig:timings-2d-map}}
  \hfill
  \subfigure[Three-dimensional case, module \code{Map}, CPU]{
    \includegraphics[width=0.475\linewidth]{images/time-3d-map.png}
    \label{fig:timings-3d-map}}
  \vskip\baselineskip
  \subfigure[Two-dimensional case, module \code{Map}, GPU]{
    \includegraphics[width=0.475\linewidth]{images/time-2d-gpu.png}
    \label{fig:timings-2d-gpu}}
  \hfill
  \subfigure[Three-dimensional case, module \code{Map}, GPU]{
    \includegraphics[width=0.475\linewidth]{images/time-3d-gpu.png}
    \label{fig:timings-3d-gpu}}
  \caption[]{Execution times for different correlation functions.}
  \label{fig:timings}
\end{figure*}

@@

\appendix
\section{Algorithm of computation of lineal-phase function}
\label{linpathalg}
The algorithm below calculates $L_2^{(phase)}$ correlation function for
one-dimensional array $a$. The result is an array of $L_2$ values for
correlation lengths from $0$ to $maxlen-1$. If $periodic$ is true (resp. false)
then periodic (resp. zero padding) boundary conditions are used. The entry point of the
algorithm is the function L2. The efficiency of the proposed algorithm is $O(n)$
where $n$ is the length of the array.

>>>>

\documentclass[reprint,amsmath,amssymb,aps,pre]{revtex4-1}
% Latex
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
% Xelatex
%\usepackage{polyglossia}
%\usepackage{fontspec}
%\setdefaultlanguage{english}
%\setmainfont{DejaVu Sans}
%\setsansfont{DejaVu Sans}
%\setmonofont{DejaVu Sans Mono}
\usepackage{bm}
\usepackage{cleveref}
\usepackage{xcolor}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{subcaption}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=.3\textwidth]{images/corr.png}
  \caption[]{{\small WTF?}}
  \label{fig:functions}
\end{figure}

@@

Correlation functions can be computed using closed walls (CW) and periodic
boundary conditions. \textbf{TODO: Describe them}.

@@

The probability $p(\bm{r})$ can be expressed as
\begin{equation}\label{eq:probability}
  p(\bm{r}) = \frac{\text{Number of successful trials $S(\bm{r})$}}
  {\text{Total number of trials $N(\bm{r})$}}
\end{equation}

@@

If $\chi(\bm{x_1}, \bm{x_2})$ can be expressed as
$\chi(\bm{x_1}, \bm{x_2}) = \chi_1(\bm{x_1})\chi_2(\bm{x_2})$ (what we call
\code{SeparableIndicator} in our package) we use the following algorithm to
calculate the array \code{success} for one-dimensional slice $a$:
\begin{algorithmic}[1]
  \Procedure{s2sep}{$a, \chi_1, \chi_2, periodic$}
  \If{$!periodic$}
    \State $a \gets a \odot zeros(length(a))$ \Comment{$\odot$
      denotes concatenation and $zeros(n)$ returns an array with $n$ zeros.}
  \EndIf
  \State $a_1 \gets \chi_1(a)$ \Comment{apply $\chi_1$ element-wise to $a$}
  \State $a_2 \gets \chi_2(a)$ \Comment{apply $\chi_2$ element-wise to $a$}
  \State $f_1 \gets rfft(a_1)$ \Comment{$rfft$ is fast Fourier transform for
    real-valued data}
  \State $f_2 \gets rfft(a_2)$
  \State $f_{S_2} \gets f_1 \cdot \overline{f_2}$ \Comment{multiply element-wise}
  \State $success \gets irfft(f_{S_2})$ \Comment{$irfft$ is inverse of $rfft$}
  \State \textbf{return} $success$
  \EndProcedure
\end{algorithmic}

If $\chi(\bm{x})$ cannot be expressed in this way (what we call
\code{InseparableIndicator}) the algorithm is slightly different:
\begin{algorithmic}[1]
  \Procedure{s2insep}{$a, \chi, periodic$}
    \State $len \gets length(a)$
    \If{$periodic$}
      \State $a \gets a \odot a$
    \EndIf
    \State $success \gets zeros(len)$
    \For{$shift \gets 0,len-1$}
      \State $a_2 \gets subseq(a, shift)$
      \Comment{$subseq$ returns subsequence of $a$ from the $shift$-th element
        to the end.}
      \State $success[shift] \gets sum(map(\chi, a, a_2))$
    \EndFor
    \State \textbf{return} $success$
  \EndProcedure
\end{algorithmic}

@@

The algorithm for calculation of the array \code{success} for
$C_2^{(i)}(\bm{r})$ for one-dimensional slice $a$ is straightforward:
\begin{algorithmic}[1]
  \Procedure{c2}{$a, phase, periodic$}
  \State $a_1 \gets x = phase$ \Comment{$a$ is compared with $phase$
    element-wise}
  \State $c \gets clusterize(a_1)$
  \Comment{$clusterize(x)$ selects clusters from the array
    $x$. Each cluster gets its own unique integer identificator. The
    identificator $0$ is reserved for the void phase.}
  \State $\chi(x,y) \gets (x = y \ne 0)$
  \State \textbf{return} $s2insep(c, \chi, periodic)$
  \EndProcedure
\end{algorithmic}

@@

\section{The map approach.}
\label{mapsec}

In this approach we actively use fast Fourier transform (FFT)
to compute some correlation function ($S_2, F_{ss}, F_{sv}, C_2$)
in all directions.
The FFT algorithm let us compute auto and cross-correlation fast,
especially on GPU.

Lineal-path function doesn't have FFT realization,
therefore, we use different approach.

\subsection{Lineal-path function}

High-level algorithm:
\begin{enumerate}
  \item Chose a set of directions
  \item for a direction in the set
    \begin{enumerate}
      \item align the image along the direction
      \item scan the aligned image
    \end{enumerate}
\end{enumerate}

Align and rotate - simple operations with linear computational complexity.
Therefore, the complexity of the entire algorithm depends linearly 
on the number of directions and on the size of the image.

\subsection{Correlation map}

\subsubsection{Auto- and cross-correlation using FFT}

All subsequent functions are auto- or cross-correlations.
Therefore, here we describe a method to compute them using FFT.

\begin{algorithmic}[1]
  \Procedure{cross\_correlation}{$a, b$}
    \State $A \gets rfft(a)$\Comment{Real fast Fourie transform}
    \State $B \gets rfft(b)$
    \ForAll{$i \in CartesianIndices(A)$}
      \State $C[i] \gets conj(A[i]) \cdot B[i]$
    \EndFor
    \State $c = irfft(C)$ 
    \Comment{inverse rfft}
    \State \textbf{return} $c$
  \EndProcedure
\end{algorithmic}

One can use cross-correlation to compute auto-correlation,
but special function would permorm better.

\begin{algorithmic}[1]
  \Procedure{auto\_correlation}{$a$}
    \State $A \gets rfft(a)$
    \ForAll{$i \in CartesianIndices(A)$}
      \State $C[i] \gets |A[i]|^2$
    \EndFor
    \State $c = irfft(C)$ 
    \State \textbf{return} $c$
  \EndProcedure
\end{algorithmic}

CartesianIndices --- all indices of multidimensional array,
therefore, the for loop describe elementwise operations
and have linear computational complexity $O(n)$.
For both functions computational complexity is $O(n\ log(n))$.

It is worth noting, that the FFT approach imply
periodic boundary conditions by design.

\subsubsection{Total number of trials}

$auto\_correlation$ and $cross\_correlation$ return non-negative number
of success trials \cref{eq:probability}.
To compute probability we need to count total number of trials.
This value depends on boundary conditions.
We realize two types: periodic and non-periodic (flat).

\begin{algorithmic}[1]
  \Procedure{total}{$a, periodic$}
    \State $m, n \gets size(a)$
    \If{periodic}
      \ForAll{$i \in range(0, m - 1)$}
        \ForAll{$j \in range(0, n - 1)$}
          \State $total[i, j] \gets m \cdot n$
        \EndFor
      \EndFor
    \Else
      \ForAll{$i \in range(-m + 1, m - 1)$}
        \ForAll{$j \in range(-n + 1, n - 1)$}
          \State $i' \gets mod(i, 2m - 1)$
          \State $j' \gets mod(j, 2n - 1)$
          \State $total[i', j'] \gets |m - i| \cdot |n - j|$
        \EndFor
      \EndFor
    \EndIf
    \State \textbf{return} $total$
  \EndProcedure
\end{algorithmic}

Described function is for 2D images,
for 1D and 3D images realization is similar.

if n is the number of pixels, then computational complexity
of this funciton is $O(n)$.

\subsubsection{Two-point funciton}

We realize two boundary conditions: periodic and non-periodic (flat).
For non-periodic boundary conditions one need to double image size
by padding new values with zeros.
Thus, if original image is $(m \times n)$,
then expanded image will have size $(2m - 1 \times 2n - 1)$.

\begin{algorithmic}[1]
  \Procedure{s2\_map}{$a, phase, periodic$}
    \State $a_1 \gets a = phase$
    \If{periodic}
      \State $a_2 \gets a_1$
    \Else
      \State $a_2 \gets expand(a_1)$\Comment{double image size by padding new values with zeros}
    \EndIf

    \State $success \gets AUTO\_CORRELATION(a_2)$

    \State $total \gets TOTAL(a, periodic)$

    \ForAll{$i \in CartesianIndices(success)$}
      \State $s2[i] \gets success[i]\ /\ total[i]$
    \EndFor

    \State \textbf{return} $s2$
  \EndProcedure
\end{algorithmic}

if n is the number of pixels, then computational complexity
of this funciton is $O(n\ log(n))$.

\subsubsection{Phase cross-correlation funciton}

\begin{algorithmic}[1]
  \Procedure{cc\_map}{$a, b, periodic$}
    \If{periodic}
      \State $a_1 \gets a$
      \State $b_1 \gets b$
    \Else
      \State $a_1 \gets expand(a)$
      \State $b_1 \gets expand(b)$
    \EndIf

    \State $success \gets CROSS\_CORRELATION(a_1, b_1)$

    \State $total \gets TOTAL(a, periodic)$

    \ForAll{$i \in CartesianIndices(success)$}
      \State $cc[i] \gets success[i]\ /\ total[i]$
    \EndFor

    \State \textbf{return} $cc$
  \EndProcedure
\end{algorithmic}

if n is the number of pixels, then computational complexity
of this funciton is $O(n\ log(n))$.

\subsection{Surface-surface function}

Computation of surface-surface function consists of two steps 
\textcolor{red}{[link to surface paper]}:
\begin{enumerate}
  \item extracting surface $M = \nabla I(a)$
  \item computing $F_{ss} = S_2(M)$
\end{enumerate}

For surface extraction we use `imgradient' function
and Sobel kernel as default kernel, 
both defined in ImageFiltering.jl package \textcolor{red}{[add link]}.

\begin{algorithmic}[1]
  \Procedure{surf\_surf}{$a, phase, periodic, kernel$}
    \State $a_1 \gets a = phase$
    \State $M \gets imgradient(a_1, kernel)$
    \State $Fss = S2\_MAP(M, periodic)$
    \State \textbf{return} $Fss$
  \EndProcedure
\end{algorithmic}

if n is the number of pixels, then computational complexity
of this funciton is $O(n\ log(n))$.

\subsection{Surface-void function}

Computation of surface-void function consists of three steps 
\textcolor{red}{[link to surface paper]}:
\begin{enumerate}
  \item extracting surface $M = \nabla I(a)$
  \item extracting void phase $V = I_v(a)$
  \item computing $F_{sv} = cross\_correlation(M, V)$
\end{enumerate}

\begin{algorithmic}[1]
  \Procedure{surf\_void}{$a, phase, periodic, kernel$}
    \State $a_1 \gets a = phase$
    \State $M \gets imgradient(a_1, kernel)$
    \State $V \gets a = 0$
    \State $Fsv = CC\_MAP(M, V, periodic)$
    \State \textbf{return} $Fsv$
  \EndProcedure
\end{algorithmic}

if n is the number of pixels, then computational complexity
of this funciton is $O(n\ log(n))$.

\subsection{Cluster funciton}

Computation of cluster function consists of three steps
\begin{enumerate}
  \item label different clusters with distinct integers
  \item compute $S_2$ map for every cluster
  \item add up all maps
\end{enumerate}

For labeling we use `label\_components' function from
Images.jl package \textcolor{red}{[add link]}.

\begin{algorithmic}[1]
  \Procedure{cluster\_map}{$a, phase, periodic$}
    \State $a_1 \gets a = phase$
    \State $labeled\_a \gets label\_components(a_1, periodic)$
    \State $number\_of\_labels = maximum(labeled\_a)$
    \ForAll{$label \in range(1, number\_of\_labels)$}
      \State $b \gets labeled\_a = label$
      \If{label = 1}
        \State $c2 \gets S2\_MAP(b, periodic)$
      \Else
        \State $c2 \gets c2 + S2\_MAP(b, periodic)$
      \EndIf
    \EndFor
    \State \textbf{return} $c2$
  \EndProcedure
\end{algorithmic}

if n is the number of pixels and N is the number of clusters, then computational complexity
of this funciton is $O(N\ n\ log(n))$.

@@

\onecolumngrid
\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/l2-2d.png}
    \caption[]{{\small Lineal-path correlation function}}
    \label{fig:l2-2d}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/s2-2d.png}
    \caption[]{{\small Two-point correlation function}}
    \label{fig:s2-2d}
  \end{subfigure}
  \vskip\baselineskip
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/ss-2d.png}
    \caption[]{{\small Surface-surface correlation function}}
    \label{fig:ss-2d}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/sv-2d.png}
    \caption[]{{\small Surface-void correlation function}}
    \label{fig:sv-2d}
  \end{subfigure}
  \caption[]{\small A comparison of calculated values of $L_2(r)$, $S_2(r)$,
    $F_{ss}(r)$ and $F_{sv}(r)$ with theoretical values for overlapping disks.}
  \label{fig:2d}
\end{figure*}

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/l2-3d.png}
    \caption[]{{\small Lineal-path correlation function}}
    \label{fig:l2-3d}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/s2-3d.png}
    \caption[]{{\small Two-point correlation function}}
    \label{fig:s3-3d}
  \end{subfigure}
  \vskip\baselineskip
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/ss-3d.png}
    \caption[]{{\small Surface-surface correlation function}}
    \label{fig:ss-3d}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/sv-3d.png}
    \caption[]{{\small Surface-void correlation function}}
    \label{fig:sv-3d}
  \end{subfigure}
  \caption[]{\small A comparison of calculated values of $L_2(r)$, $S_2(r)$,
    $F_{ss}(r)$ and $F_{sv}(r)$ with theoretical values for overlapping balls.}
  \label{fig:3d}
\end{figure*}

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/ps-2d.png}
    \caption[]{{\small Pore size correlation function (2D)}}
    \label{fig:ps-2d}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/cl-2d.png}
    \caption[]{{\small Chord length correlation function (2D)}}
    \label{fig:cl-2d}
  \end{subfigure}
  \vskip\baselineskip
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/ps-3d.png}
    \caption[]{{\small Pore size correlation function (3D)}}
    \label{fig:ps-3d}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/cl-3d.png}
    \caption[]{{\small Chord length correlation function (3D)}}
    \label{fig:cl-3d}
  \end{subfigure}
  \caption[]{\small A comparison of calculated values of $P(r)$ and $p(r)$
    with theoretical values for overlapping disks (\cref{fig:ps-2d} and
    \cref{fig:cl-2d}) and balls(\cref{fig:ps-3d} and \cref{fig:cl-3d}).}
  \label{fig:pscl}
\end{figure*}
\twocolumngrid

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=.3\textwidth]{images/timing-image.png}
  \caption[]{{\small An example of 2D image used in timing measurements}}
  \label{fig:valuenoise}
\end{figure}

\onecolumngrid
\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/time-2d.png}
    \caption[]{{\small Two-dimensional case, module \code{Directional}}}
    \label{fig:timings-2d}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/time-3d.png}
    \caption[]{{\small Three-dimensional case, module \code{Directional}}}
    \label{fig:timings-3d}
  \end{subfigure}
  \vskip\baselineskip
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/time-2d-map.png}
    \caption[]{{\small Two-dimensional case, module \code{Map}, CPU}}
    \label{fig:timings-2d-map}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/time-3d-map.png}
    \caption[]{{\small Three-dimensional case, module \code{Map}, CPU}}
    \label{fig:timings-3d-map}
  \end{subfigure}
  \vskip\baselineskip
    \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/time-2d-gpu.png}
    \caption[]{{\small Two-dimensional case, module \code{Map}, GPU}}
    \label{fig:timings-2d-gpu}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/time-3d-gpu.png}
    \caption[]{{\small Three-dimensional case, module \code{Map}, GPU}}
    \label{fig:timings-3d-gpu}
  \end{subfigure}
  \caption[]{\small Execution times for different correlation functions.}
  \label{fig:timings}
\end{figure*}
\twocolumngrid

@@

\appendix
\section{Algorithm of computation of lineal-phase function}
\label{linpathalg}
The algorithm below calculates $L_2^{(phase)}$ correlation function for
one-dimensional array $a$. The result is an array of $L_2$ values for
correlation lengths from $0$ to $maxlen-1$. If $periodic$ is true (resp. false)
then periodic (resp. CW) boundary conditions are used. The entry point of the
algorithm is the function L2. The efficiency of the proposed algorithm is $O(n)$
where $n$ is the length of the array.
